---
title: "GR5065 Assignment 1"
author: "Nashrah Ahmed - na2729"
date: "2/1/18"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(123)

```

## Question 1
### Part 1.1
```{r}
# Scenario 1:
# roll 1 (strike)

F <-function(x){  
  stopifnot(is.numeric(x), all(x == as.integer(x)))  
  sqrt_5 <- sqrt(5) # defined once, used twice  
  golden_ratio <-(1+ sqrt_5)/2
  return(round(golden_ratio ^(x +1)/ sqrt_5))
}
# probability of knocking down x out of n pins
Pr <- function(x, n = 10) 
  return(ifelse(x > n,0, F(x)/(-1+ F(n +2))))

Omega <- 10
sum(Pr(Omega))

# roll 2 (strike) assuming no interdependence given new set of 10 pins
F <-function(x){  
  stopifnot(is.numeric(x), all(x == as.integer(x)))  
  sqrt_5 <- sqrt(5) # defined once, used twice  
  golden_ratio <-(1+ sqrt_5)/2
  return(round(golden_ratio ^(x +1)/ sqrt_5))
}
# probability of knocking down x out of 10 pins
Pr <- function(x, n = 10) 
  return(ifelse(x > n,0, F(x)/(-1+ F(n +2))))

Omega <- 10
sum(Pr(Omega))

# roll 2, x number of pins assuming strike on 1st roll
F <-function(x){  
  stopifnot(is.numeric(x), all(x == as.integer(x)))  
  sqrt_5 <- sqrt(5) # defined once, used twice  
  golden_ratio <-(1+ sqrt_5)/2
  return(round(golden_ratio ^(x +1)/ sqrt_5))
}

# probability of knocking down x out of 10 pins
Pr <- function(x, n = 10) 
  return(ifelse(x > n,0, F(x)/(-1+ F(n +2))))

Omega <- 0:10
sum(Pr(Omega))

x3 <- sample(Omega, size = 1, prob = Pr(Omega)) # realization of random variable

# roll 3 (strike) assuming no interdependence given new set of 10 pins
Pr <- function(x, n = 10) 
  return(ifelse(x > n,0, F(x)/(-1+ F(n +2))))

Omega <- 0:10
sum(Pr(Omega))

x4 <- sample(Omega, size = 1, prob = Pr(Omega)) # realization of random variable

# or conditional prob based on # of pins knocked down in roll 2

sum(Omega * Pr(Omega, n = 10 - x4))

# Conditional expectation (1) = 

```
### Part 1.2
```{r}

# Scenaroio 2:
# roll 1 (knock down 0 to 9 pins)
F <-function(x){  
  stopifnot(is.numeric(x), all(x == as.integer(x)))  
  sqrt_5 <- sqrt(5) # defined once, used twice  
  golden_ratio <- (1+ sqrt_5)/2
  return(round(golden_ratio ^(x +1)/ sqrt_5))
}
# probability of knocking down x out of n pins
Pr <- function(x, n = 10) 
  return(ifelse(x > n,0, F(x)/(-1+ F(n +2))))

Omega <- 0:9
sum(Pr(Omega))

x5 <- sample(Omega, size = 1, prob = Pr(Omega)) # realization of random variable

# roll 2 knock down remainder (how to ensure stike??)

n = 10 - x5

sum(Omega * Pr(n, n))

# roll 3 probability of knocking down x out of 10 pins

F <-function(x){  
  stopifnot(is.numeric(x), all(x == as.integer(x)))  
  sqrt_5 <- sqrt(5) # defined once, used twice  
  golden_ratio <-(1+ sqrt_5)/2
  return(round(golden_ratio ^(x +1)/ sqrt_5))
}

# probability of knocking down x out of 10 pins

Pr <- function(x, n = 10) 
  return(ifelse(x > n,0, F(x)/(-1+ F(n +2))))

Omega <- 0:10
sum(Pr(Omega))

x6 <- sample(Omega, size = 1, prob = Pr(Omega)) # realization of random variable

# Conditional expectation (2) = 

```

### Part 1.3
```{r}

# Scenario 3
# roll 1 (knock down 0 to 9 pins)
F <-function(x){  
  stopifnot(is.numeric(x), all(x == as.integer(x)))  
  sqrt_5 <- sqrt(5) # defined once, used twice  
  golden_ratio <- (1+ sqrt_5)/2
  return(round(golden_ratio ^(x +1)/ sqrt_5))
}
# probability of knocking down x out of n pins
Pr <- function(x, n = 10) 
  return(ifelse(x > n,0, F(x)/(-1+ F(n +2))))

Omega <- 0:9
sum(Pr(Omega))

x7 <- sample(Omega, size = 1, prob = Pr(Omega)) # realization of random variable

# roll 2 knock down with at least one remaining ??

n = 10 - x7

sum(Omega * Pr(n-1, n))

# Conditional expectation (3) = 

## Total Expectation

```
Realized my attempts above were a very long-winded/incorrect way of calculating after seeing the solutions.

## Question 2
### Part 2.1
What is the probability of being dealt any two Aces?

Probability of getting an Ace and another Ace is equal to probability of getting an Ace times the probability of getting the 2nd Ace haven already recieved an Ace. 

$$ 4/52 * 3/51 = ~ 0.0045 $$
What is the probability of being dealt any pair of cards that have the same value?

$$ 13/221 = ~ 0.059 $$

### Part 2.2
Betting first

Betting first seems to offer less optionality in terms of strategy as the 1st player has a limited amount of information relative to others. The only real information the 1st player has at that time is his/her own deck which can either be perceived as good or bad. The players that go after have information from past bets and folds.

### Part 2.3

### Part 2.4
Pair of pairs

Probability of getting an Ace and another Ace for two players is equal to the probability of getting an Ace times the probability of getting the 2nd Ace haven already dealt an Ace times the the probability of getting the 3rd Ace haven already dealt 2 Aces times the probability of picking a 4th Ace haven already dealt 3 Aces.

$$ 4/52 * 3/51 * 2/50 * 1/49 = ~ 0.00000369 $$

### Part 2.5
Bluffing

Bluffing can be advantageous or a successful approach only if the opponent is mislead by your buffling (or tricked into believing that you have a better hand than you do).

### Part 2.6 

### Part 2.7
Kings

Probability that Katz has 2 Kings is equal to the probability of having a King times the probability of having a secong King given 1 King present. 

$$ Pr(K1) × Pr(K2|K1) = 4/6 × 3/5 = 0.4 $$

### Part 2.8

### Part 2.9
Side commentary
The comment that Antonio makes can be considered a Bayesian statement if its indeed based on his beliefs of how many Aces exsist in the deck and his acknowledgment of how many Aces he has seen.

### Part 2.10 

### Part 2.11

## Question 3
```{r}
# Bernoulli and Poisson

S <- 1000
mean(4 == rbinom(S, size = rpois(S, sqrt(7)), prob = 2 / 3))

```

## Question 4
Frequentism

An example where one could consider the frequentist approach to be appropriate is a project I was exposed to by a NLP professor last semester. The project used NLP techniques to collect data from police officers in Ohio to measure the level of "politeness" and "rudeness" based on the tone of the officers when addressing people from various ethnic backgrounds, specifically white people vs black people.This study used regression to determine the likelihood an officer is to behave x way based on the way he/she had behaved in the past, taking into account several factors (ie. type of car, type of crime, etc,). The sample data seemed narrow and consistant enough to support an inference approach in this case. 
